---
title: "Final"
date: "2026-05-12"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
#packages
library(dplyr)
library(tidyverse)
library(estimatr)
library(stargazer)
```
```{r}
#Setting working directory to where the csv file is saved
setwd("/Users/jovanakovac/Documents/Eco590/Data_Analysis_Python_R/final-project-jovanakovac")

world_data <- read.csv("world_data.csv")
options(scipen = 100)
```
```{r}
#Regression 1: Does log GDP per capita predict life expectancy,
#Controlling for unemployment rate and health expenditure
#Log GDP is used because the relationship between income and life expectancy is not linear

ols <- lm(life_expectancy ~ log_gdp + unemployment_rate + health_expenditure, data = world_data)
summary(ols)

print(colnames(world_data))
```
```{r}
#Regression 2: Filtering for countries with above average GDP per capita
#Isolating wealthier countries to see if the relationship still holds within this group
world_high <- world_data[world_data$log_gdp > mean(world_data$log_gdp, na.rm = TRUE),]
ols_high <- lm(life_expectancy ~ log_gdp + unemployment_rate + health_expenditure + log_gdp:health_expenditure, data = world_high)
summary(ols_high)
```

```{r}
ols_p <- lm(life_expectancy ~ log_gdp + unemployment_rate + health_expenditure, data = world_data)
summary(ols_p)
```
```{r}
#Displaying all regressions together
stargazer(ols, ols_high, type="text",
          title = "Life Expectancy Regression Results",
          dep.var.labels = c("Life Expectancy", "Life Expectancy (High GDP)"),
          covariate.labels = c("Log GDP per Capita", "Unemployment Rate", "Health Expenditure", "Log GDP x Health Expenditure"))
```

## Interpretation

Regression 1 shows that holding unemployment rate and health expenditure constant, a one percent increase in GDP per capita is associated with 1.555 additional years of life expectancy. This result is statistically significant at the 1% level. Unemployment rate is not statistically significant. Health expenditure is positive and statistically significant.

Regression 2 is the preferred regression. It filters for countries with above-average GDP per capita and adds an interaction term between log GDP and health expenditure. The interaction term is positive and significant at the 5% level, suggesting that in wealthier countries, the combined effect of GDP and health spending is stronger. This is the preferred regression.
